Remote sensing riverine gas exchange and carbon efflux in ungauged rivers from SWOT observations

### 2022-06-08 docs/intial_drafts/manuscript_v11.Rmd
first round of review, setting up for rerun on HPC

River $CO_2$ evasion is increasingly better constrained and is clearly a critical component of the global carbon cycle. Equation 1 represents this riverine flux given $\delta CO_2$ (the air-water flux between the water $CO_{2_{water}}$ and the air $CO_{2_{air}}$) and the gas exchange velocity *k*. Consult Appendix A for variable nomenclature used throughout this study.

@@

Figure 2 shows that Equation 7 explains `r round(modelsWIDE[modelsWIDE$name == 'reynolds-eD',]$r2, 2) * 100`% of variation in observed $k_{600}$ in hydraulically-wide rivers and accurately captures the scaling dynamics of observed $k_{600}$ as well. The value for $\beta_1$ is 62.8, which is similar the fitted parameters in other $Re_*$ and $U_*$ models for river *k* [@wangPhysicallyBasedScaling2021]. While the individual residuals can be quite large, the overall scaling of $k_{600}$ with river hydraulics is strongly captured, remembering that Figure 2 is plotted in logarithmic space. Compared to other combinations of Equations 3-6 (Figure S1), there is also a better fit with the estimates thanks to the addition of the Reynold's number scaling for low-turbulent flows (as expected). The success of this model in hydraulically-wide channels provides us with a strong physical-model for gas evasion built with SWOT in mind. The river hydraulics terms in Equation 7 ($\bar{U}$, $H$, and $S$) can either be directly measured or reasonably inferred from SWOT measurements, effectively opening the door for remotely sensing the gas exchange velocity.

@@

We validate BIKER under two different 'error scenarios'. While SWOT will provide river surface measurements of novel quality and resolution, as with all remote sensing products there are expected errors that will be implicit in these measurements. Here, we validate BIKER under a 'no-measurement-error' scenario that reflects an unrealistic measurement as if SWOT has perfect accuracy and precision: we use the hydraulic model output directly as a first test of algorithm validity. `r nrow(results_k600_rivs[results_k600_rivs$errFlag == 1,])` of these rivers are then additionally validated under a 'measurement-error' scenario that more closely mimics expected SWOT by adding realistic radar errors and sampling along the satellite's future ground track. SWOT river error modeling was developed by @durandHowWillRadar2020 and then applied to these `r nrow(results_k600_rivs[results_k600_rivs$errFlag == 1,])` rivers by @frassonExploringFactorsControlling2021. This error modelling is non-trivial and computationally expensive, and thus we rely on these previously published data.

@@

$$\left[log(62.82) + \frac{7}{16}log(g) + \frac{9}{16}log(S_{i,t}) - \frac{35}{48}log(W_{i,t})\right] \sim normal\left(\left[log(k_{600_t})+\frac{1}{4}log(n_i)-\frac{35}{48}log(A_0_i+dA_{i,t})\right], \sigma^2_{k_{600}}\right) \mathbf{(10)}$$ Equations 8 and 10 also necessitate that we specify prior distributions for the parameters $A_0_i$, $k_{600_t}$, and $n_i$. Prior distributions formalize the a priori estimates and uncertainties for the non-remotely-sensed terms. More intuitively, BIKER priors represent our 'prior river knowledge' of what $A_0_i$, $k_{600_t}$, and $n_i$ probably are for some river since they cannot be directly remotely sensed. This is analogous to the 'empirical Bayes approach' to Bayesian inference [@hoffFirstCourseBayesian2009]. Our goal in prior specification was to rely on absolutely no in situ information such that we could run this method on any river on Earth solely using SWOT observations. In theory, more informed priors via various a priori information about a specific river will improve BIKER performance, but here we chose to test the fully generalized algorithm. Therefore, the validation presented here is a 'worst-case scenario', wherein BIKER performance should improve with better prior information on the river. In that context, we used a variation of the prior specification method developed by @brinkerhoffConstrainingRemoteRiver2020, who developed 'geomorphic river types' with distinct prior sets for $A_0_i$ and $n_i$. These priors are assigned to a river solely using SWOT data *W* and *S*, therefore meeting our needs for complete global implementability. Prior assignment for $k_{600_t}$ was developed similarly (all prior specifications are elaborated on in Text S3).

@@

Figure 7 compares the annual carbon emissions rate (via $F_{CO_2}$) from the rivers using BIKER posterior means and the three stream gauge-based HG models. Surprisingly, BIKER outperforms the gauge-based approaches, nearly correctly inferring the annual carbon emissions rate (`r round(results_fco2_bulk[results_fco2_bulk$key == 'FCO2_BIKER',]$yearlyCFux, 2)` Tg-C/yr for BIKER versus `r round(results_fco2_bulk[results_fco2_bulk$key == 'FCO2_obs',]$yearlyCFux, 2)` Tg-C/yr observed). The three HG models overestimate this emissions rate: `r round(results_fco2_bulk[results_fco2_bulk$key == 'FCO2_raymond2013',]$yearlyCFux, 2)`, `r round(results_fco2_bulk[results_fco2_bulk$key == 'FCO2_raymond2012',]$yearlyCFux, 2)`, and `r round(results_fco2_bulk[results_fco2_bulk$key == 'FCO2_brinkerhoff2019',]$yearlyCFux, 2)` for 'Raymond 2013', 'Raymond 2012', and 'Brinkerhoff 2019' respectively. 'Raymond 2012' falls within the BIKER CIs and is reasonably close to the observed value, while 'Raymond 2013' overestimates the emission rate. 'Brinkerhoff 2019''s lower CI is comparable to BIKER's higher CI. BIKER's relatively stronger performance than the in situ models is elaborated on in Section 4.3. Finally, BIKER's uncertainity is on par with the in situ technique, despite being obtained solely from SWOT data. Taken in aggregate, BIKER provides a strong upscaling estimate of the annual carbon emission rate for the rivers and is either comparable or better than established in situ techniques (Figure 7).

@@

First, Figure 7 directly implies that BIKER will be useful when coupled with large-scale $CO_{2_{water}}$ models, provided these models are accurate. The models would give time and space varying gas exchange. @liuImportanceHydrologyRouting2022 and @saccardiImprovingPredictionsStream2021 each propose models that robustly predict reach-scale dissolved $CO_2$ concentrations using two different approaches [machine learning- @liuImportanceHydrologyRouting2022 and process-based reactive transport modeling- @saccardiImprovingPredictionsStream2021], but both models yield $CO_2$ estimates that would be spatially and temporally consistent with BIKER's output. Our promising results suggest that BIKER could provide additional (and directly inferred) measurements of $k_{CO_2}$ to these models, thereby better informing model results through direct observations. This is likely to be accomplished via data assimilation which has proven useful in using remotly-sensed discharge to improve streamflow routing models [@fengRecentChangesArctic2021; @ishitsukaCombiningOpticalRemote2020], and of which the @saccardiImprovingPredictionsStream2021 $CO_2$ model takes a similar form.

>>>>

River $CO_2$ evasion is increasingly better constrained and is clearly a critical component of the global carbon cycle. Equation 1 represents this riverine flux given $\delta CO_2$ (the air-water difference between the water $CO_{2_{water}}$ and the air $CO_{2_{air}}$) and the gas exchange velocity *k*. Consult Appendix A for variable nomenclature used throughout this study.

@@

Figure 2 shows that Equation 7 explains `r round(modelsWIDE[modelsWIDE$name == 'reynolds-eD',]$r2, 2) * 100`% of variation in observed $k_{600}$ in hydraulically-wide rivers and accurately captures the scaling dynamics of observed $k_{600}$ as well. The value for $\beta_1$ is 62.8, which is similar to the fitted parameters in other $Re_*$ and $U_*$ models for river *k* [@wangPhysicallyBasedScaling2021]. While the individual residuals can be quite large, the overall scaling of $k_{600}$ with river hydraulics is strongly captured, remembering that Figure 2 is plotted in logarithmic space. Compared to other combinations of Equations 3-6 (Figure S1), there is also a better fit with the estimates thanks to the addition of the Reynold's number scaling for low-turbulent flows (as expected). The success of this model in hydraulically-wide channels provides us with a strong physical-model for gas evasion built with SWOT in mind. The river hydraulics terms in Equation 7 ($\bar{U}$, $H$, and $S$) can either be directly measured or reasonably inferred from SWOT measurements, effectively opening the door for remotely sensing the gas exchange velocity.

@@

We validate BIKER under two different 'error scenarios' (section 3.2.2). While SWOT will provide river surface measurements of novel quality and resolution, as with all remote sensing products there are expected errors that will be implicit in these measurements. Here, we validate BIKER under a 'no-measurement-error' scenario that reflects an unrealistic measurement as if SWOT has perfect accuracy and precision: we use the hydraulic model output directly as a first test of algorithm validity. `r nrow(results_k600_rivs[results_k600_rivs$errFlag == 1,])` of these rivers are then additionally validated under a 'measurement-error' scenario that more closely mimics expected SWOT by adding realistic radar errors and sampling along the satellite's future ground track. SWOT river error modeling was developed by @durandHowWillRadar2020 and then applied to these `r nrow(results_k600_rivs[results_k600_rivs$errFlag == 1,])` rivers by @frassonExploringFactorsControlling2021. This error modelling is non-trivial and computationally expensive, and thus we rely on these previously published data.

@@

$$\left[log(62.82) + \frac{7}{16}log(g) + \frac{9}{16}log(S_{i,t}) - \frac{35}{48}log(W_{i,t})\right] \sim \mathcal{N}\left(\left[log(k_{600_t})+\frac{1}{4}log(n_i)-\frac{35}{48}log(A_0_i+dA_{i,t})\right], \sigma^2_{k_{600}}\right) \mathbf{(10)}$$ Equations 8 and 10 also necessitate that we specify prior distributions for the parameters $A_0_i$, $k_{600_t}$, and $n_i$. Prior distributions formalize the a priori estimates and uncertainties for the non-remotely-sensed terms. More intuitively, BIKER priors represent our 'prior river knowledge' of what $A_0_i$, $k_{600_t}$, and $n_i$ probably are for some river since they cannot be directly remotely sensed. This is analogous to the 'empirical Bayes approach' to Bayesian inference [@hoffFirstCourseBayesian2009]. Our goal in prior specification was to rely on absolutely no in situ information such that we could run this method on any river on Earth solely using SWOT observations. In theory, more informed priors via various a priori information about a specific river will improve BIKER performance, but here we chose to test the fully generalized algorithm. Therefore, the validation presented here is a 'worst-case scenario', wherein BIKER performance should improve with better prior information on the river. In that context, we used a variation of the prior specification method developed by @brinkerhoffConstrainingRemoteRiver2020, who developed 'geomorphic river types' with distinct prior sets for $A_0_i$ and $n_i$. These priors are assigned to a river solely using SWOT data *W* and *S*, therefore meeting our needs for complete global implementability. Prior assignment for $k_{600_t}$ was developed similarly (all prior specifications are elaborated on in Text S3).

@@

Figure 7 compares the annual carbon emissions rate (via $F_{CO_2}$) from the rivers using BIKER posterior means and the three stream gauge-based HG models. Surprisingly, BIKER outperforms the gauge-based approaches, nearly correctly inferring the annual carbon emissions rate (`r round(results_fco2_bulk[results_fco2_bulk$key == 'FCO2_BIKER',]$yearlyCFux, 2)` Tg-C/yr for BIKER versus `r round(results_fco2_bulk[results_fco2_bulk$key == 'FCO2_obs',]$yearlyCFux, 2)` Tg-C/yr observed). The three HG models overestimate this emissions rate: `r round(results_fco2_bulk[results_fco2_bulk$key == 'FCO2_raymond2013',]$yearlyCFux, 2)`, `r round(results_fco2_bulk[results_fco2_bulk$key == 'FCO2_raymond2012',]$yearlyCFux, 2)`, and `r round(results_fco2_bulk[results_fco2_bulk$key == 'FCO2_brinkerhoff2019',]$yearlyCFux, 2)` for 'Raymond 2013', 'Raymond 2012', and 'Brinkerhoff 2019' respectively. 'Raymond 2012' falls within the BIKER CIs and is reasonably close to the observed value, while 'Raymond 2013' overestimates the emission rate. 'Brinkerhoff 2019''s lower CI is comparable to BIKER's higher CI. BIKER's relatively stronger performance than the in situ models is elaborated on in Section 4.3. Finally, BIKER's uncertainity is on par with the in situ technique (‘Brinkerhoff 2019’), despite being obtained solely from SWOT data. Taken in aggregate, BIKER provides a strong upscaling estimate of the annual carbon emission rate for the rivers and is either comparable or better than established in situ techniques (Figure 7).

@@

First, Figure 7 directly implies that BIKER will be useful when coupled with large-scale $CO_{2_{water}}$ models, provided these models are accurate. The models would give time and space varying gas exchange. @liuImportanceHydrologyRouting2022 and @saccardiImprovingPredictionsStream2021 each propose models that robustly predict reach-scale dissolved $CO_2$ concentrations using two different approaches- machine learning for @liuImportanceHydrologyRouting2022 and process-based reactive transport modeling for @saccardiImprovingPredictionsStream2021- but both models yield $CO_2$ estimates that would be spatially and temporally consistent with BIKER's output. Our promising results suggest that BIKER could provide additional (and directly inferred) measurements of $k_{CO_2}$ to these models, thereby better informing model results through direct observations. This is likely to be accomplished via data assimilation which has proven useful in using remotly-sensed discharge to improve streamflow routing models [@fengRecentChangesArctic2021; @ishitsukaCombiningOpticalRemote2020], and of which the @saccardiImprovingPredictionsStream2021 $CO_2$ model takes a similar form.

### 2022-06-08 docs/intial_drafts/manuscript_v3.Rmd
first round of review, setting up for rerun on HPC

##1 Introduction
**I don't like this opening paragraph** The Earth's carbon cycle moves carbon between the atmosphere, geosphere, hydrosphere, and biosphere via a complex set of interconnected biogeochemical processes. The primary connection between the terrestrial and oceanic biospheres is the global river network, which exports a significant amount of carbon from the continents to the oceans [@colePlumbingGlobalCarbon2007a], while additionally transporting and transforming sediments, nutrients, organic matter, dissolved gases, and other constituents. Organic matter and inorganic nutrients are also produced and/or consumed within rivers via ecosystem metabolism [@odumPrimaryProductionFlowing1956]. One net effect of all of this is the global river network is usually supersaturated with dissolved gases and evades substantial amounts of carbon from its water surface to the atmosphere [@colePlumbingGlobalCarbon2007a].

Current estimates of total carbon dioxide ($CO_2$) evasion to the atmosphere vary from 650-1800 Tg C/yr [@lauerwaldSpatialPatternsCO22015; @raymondGlobalCarbonDioxide2013a], with 167 Tg-C/yr coming from mountain streams alone [@horgbyUnexpectedLargeEvasion2019]. Further, in-river metabolism is increasingly understood to contribute to this flux [@hotchkissSourcesProcessesControlling2015]. Equation 1 represents this flux for any sparingly soluble gas $[M/L^2T]$ given the gas concentration gradient between the water $C_{water}$ $[M/L^3]$ and the air $C_{air}$ $[M/L^3]$ and the gas exchange velocity *k [L/T]*. Given the structure of equation 1, calculations of this flux are very sensitive to one's measurements/estimates of *k*. Note that *k* is different for different gases and water temperatures [see @hallGasExchangeStreams2020 for a thorough review of gas exchange in rivers].

@@

*k* is also vital to understanding freshwater ecology through bulk river metabolism [more specifically, gross primary production and ecological respiration- @bernhardtMetabolicRegimesFlowing2018]. The general mass-balance equation for river metabolism (*M* $[M/L^2T]$) is provided as equation 2, where *DO* is the dissolved oxygen concentration $[M/L^3]$, *D* is river depth [L], and $k_{o_2}$ is *k* for oxygen at some temperature. When viewing equations 1 and 2 in aggregate, it is clear that *k* plays a critical role in two of the core, fundamental biogeochemical processes that occur in river networks. Both are reliant on accurate and high-fidelity assessments of *k*, as well as a rigorous understanding of the physical mechanisms driving *k*.
$$\mathbf{(2)} \frac{dDO}{dt}=\frac{M}{D}+\frac{k_{o_2}}{D}(DO_{air}-DO_{water})$$
Researchers use a range of methods to measure *k*, from floating domes to eddy covariance towers to measure at a specific point in the river. They also inject tracer gases to estimate reach-scale gas exchange [@hallGasExchangeStreams2020]. These approaches are ideal but are infeasible when working in ungauged rivers or at the river network scale across thousands of rivers. So, predictive models are used to estimate *k* from readily available river geomorphology [e.g. @oconnorMechanismReaerationNatural1958; @palumboAssessingPerformanceReaeration2014; @raymondScalingGasTransfer2012a]. These models often, though not always, predict $k_{600}$ or *k* that has been normalized to a Schmidt number of 600 ($CO_2$ at 20 degrees Celsius or $O_2$ at 17.5 degrees).

@@

**I'm struggling with this paragraph, and also don't know how to write it diplomatically**
We implement this model within BIKER (using 48 as a reasonable value for $\alpha$ given both forms of equation 3) for a few reasons. First, @wangPhysicallyBasedScaling2021 suggest that an $\alpha$ of approximately 48 is uniform across rivers and streams, regardless of the training data used. This is encouraging for use within BIKER as BIKER is specifically designed to be as river-agnostic as possible and so we sought as generalized an equation as possible. Second, it yields a simple linear relationship across all rivers, regardless of their size or steepeness. This is not necessairly true of other *k* models, for example those based on *eD* (section 1). The relationship between *eD* and *k* is not necessairly linear [@raymondScalingGasTransfer2012a], suggesting poorer performance for small *k* values. Further, recent work has shown that *k* does not scale with *eD* via the same statistical parameters across all river geomorphologies [@ulsethDistinctAirWater2019a]. We further explored the relationship presentedby [@ulsethDistinctAirWater2019a] and found that the $k \sim eD$ relationship fundamentally breaks down in rivers with low *eD*, which corresponds to nearly all rivers that SWOT will observe (Figure S2). Therefore, an *eD* based predictive model is less useful in SWOT rivers. Third, we sought to explicitly predict $k_{o_2}$, as its use extends beyond constraining gas fluxes from rivers and into parameterizing river metabolism models (equation 2). We stress that most other predictive models for *k* could be implemented within BIKER. However, our goal of using BIKER on any river that is observable by SWOT necessitates that we pick an equation that is applicable across as many rivers as possible. Future work should look at using a river-specific *k* equation to pseudo-calibrate BIKER to a specific river.

@@

With the likelihood function (equation 6), prior distributions, and the $\sigma_{k_{o_2}}^2$ parameter described (section 3.2.2), a joint posterior distribution conditional on the SWOT observations is specified. To approximate this distribution, we use a Markov Chain Monte Carlo (MCMC) algorithm implemented using the Stan probabilistic programming language. Specifically, BIKER uses a Hamiltonian Monte Carlo sampler which reduces computation time relative to other sampling algorithms [@hagemannBAMBayesianAMHGManning2017a].

###3.3 Validation setup
We validate BIKER on the aforementioned 49 SWOT-simulated rivers (section 2) using daily observed $R_h$ and observed $S_h$. We also re-validate BIKER on the 17 rivers with the SWOT error model corrupting the SWOT-observables *W* and *H*. However, regardless of the validation data or SWOT error budget used, we do not have observed $k_{o_2}$ data for these rivers, and to our knowledge no field dataset of $k_{o_2}$ exists in the type of temporal and spatial frequency that SWOT (and therefore the BIKER algorithm) provides. Further, we are principally interested in BIKER's ability to reproduce $k_{o_2}$ from SWOT observations and are less concerned with the actual accuracy of the predictive model itself, which can be validated using existing datasets as has been previously done [e.g. @wangPhysicallyBasedScaling2021; @ulsethDistinctAirWater2019a; @raymondScalingGasTransfer2012a]. Therefore, we take the model outlined in section 3.1 and use that to calculate the observed $k_{o_2}$ that BIKER is validated against. This is done using equation 8, where $R_{h, obs} [L/T]$ is observed river channel area divided by observed channel width. This means that $k_{o_2, obs}$ relies on in situ river shear velocity, while BIKER is predicting shear velocity from *W* and $S_h$ alone. Therefore, we are directly exploring BIKER's ability to infer observed $U_\star$ and $k_{o_2}$ from *W* and *H* alone. It also means that, for a fair validation scheme, $\sigma_{k_{o_2}}^2$ must be set to reflect only error from our assumptions about $R_h$ and not the parameter uncertainity inherent in $\alpha$ (the 48 in equation 8). Thus, $\sigma_{k_{o_2}}^2$ is set to 0.20 for this validation as suggested in section 3.2. However once SWOT launches, it should reflect the total uncertainty calculated in section 3.2 from both sources.

@@

###3.4 Upscaling to CO2 evasion and bulk carbon efflux
It is one thing to accurately predict $k_{o_2}$, but researchers are often most interested in the actual gas fluxes from rivers and ultimately the carbon emitted from river to atmosphere. Therefore, we explore 1) the ability of BIKER to reproduce $CO_2$ evasion fluxes ($FCO_2$ $[M/L^2T]$) from these 49 rivers, and 2) the sensitivity of the carbon efflux $[M/T]$ due to the chosen method to approximate $R_h$ and ultiamtely $U_\star$.

First, we calculate $FCO_2$. To do this, we pair the biweekly 26 $CO_2$ samples (section 2) with every 11th SWOT observation by date, ignoring the timesteps beyond 26 (only ~15% of the SWOT observations were ignored here and we deemed this acceptable). We chose to sample every 11 days as this is the average sampling resolution for SWOT and the $CO_2$ data is only every 14 days. Not all of the SWOT rivers have observations for a full year, and when simulation dates were not available they were assumed to start on January 1st. We also pair the modeled $k_{CO_2}$ values (obtained from $k_{o_2}$ and equation 9) with these water-side $CO_2$ concentrations and water temperatures. In equation 9, 530 is the Schmidt number for oxygen at 20 degrees Celsius (which is the temperature at which BIKER predicts *k*) and *Sc* is the Schmidt number at one's desired temperature for some dissolved gas. Atmospheric $CO_2$ was assumed 390 uatm. The Schmidt number, used in equation 9, was calculated following @raymondScalingGasTransfer2012a and @wanninkhofRelationshipWindSpeed1992a. $FCO_2$ validation was performed using the same metrics as $k_{o_2}$ validation (Table 1).

@@

Next, we estimate bulk carbon efflux using five models for average channel depth (used to calculate $U_\star$):  BIKER's posterior means and four gauge-based HG models previously used for $CO_2$ upscaling [@horgbyUnexpectedLargeEvasion2019; @lauerwaldSpatialPatternsCO22015; @raymondGlobalCarbonDioxide2013a; @raymondScalingGasTransfer2012a]. See  for their definitions. It is worth stressing that the @lauerwaldSpatialPatternsCO22015 model is actually one of two components of the @raymondGlobalCarbonDioxide2013a model, and so was developed by the latter's authors. The names used here refer to the specific implementation used in each study. The @horgbyUnexpectedLargeEvasion2019 was developed explicitly for steep mountain streams, of which SWOT will not sample. We include it here to explore just how sensitive efflux estimates are to the HG model that is employed, noting that it will likely perform poorly. For this study, all HG methods use the in-situ discharge record while BIKER does not. This allows us to assess whether BIKER's estimates (wholly ungauged) are comparable to gauged methods (all four HG models). Finally, we express the bulk carbon efflux as the average mass flow rate of carbon (via $CO_2$ evasion) per year from the 49 rivers after accounting for total river surface area.

>>>>

##1 Introduction
**I don't like this opening paragraph** The Earth moves carbon between the atmosphere, geosphere, hydrosphere, and biosphere via a complex set of interconnected biogeochemical processes. The primary connection between the terrestrial and oceanic biospheres is the global river network, which exports a significant amount of carbon from the continents to the oceans [@colePlumbingGlobalCarbon2007a], while additionally transporting and transforming sediments, nutrients, organic matter, dissolved gases, and other constituents. Organic matter and inorganic nutrients are also produced and/or consumed within rivers via ecosystem metabolism [@odumPrimaryProductionFlowing1956]. One net effect of all of this is the global river network is usually supersaturated with dissolved gases and evades substantial amounts of carbon from its water surface to the atmosphere [@colePlumbingGlobalCarbon2007a].

Current estimates of total carbon dioxide ($CO_2$) evasion to the atmosphere vary from 650-1800 Tg C/yr [@lauerwaldSpatialPatternsCO22015; @raymondGlobalCarbonDioxide2013a], with 167 Tg-C/yr coming from mountain streams alone [@horgbyUnexpectedLargeEvasion2019]. Further, in-river metabolism is increasingly understood to contribute to this flux [@hotchkissSourcesProcessesControlling2015]. Equation 1 represents this flux for any sparingly soluble gas $[M/L^2T]$ given the gas concentration gradient between the water $C_{water}$ $[M/L^3]$ and the air $C_{air}$ $[M/L^3]$ and the gas exchange velocity *k [L/T]*. Given the structure of equation 1, calculations of this flux are highly sensitive to one's measurements/estimates of *k*. Note that *k* is different for different gases and water temperatures [see @hallGasExchangeStreams2020 for a thorough review of gas exchange in rivers].

@@

*k* is also vital to understanding freshwater ecology through bulk river metabolism [more specifically, gross primary production and ecological respiration- @bernhardtMetabolicRegimesFlowing2018]. The general mass-balance equation for river metabolism (*M* $[M/L^2T]$) is provided as equation 2, where *DO* is the dissolved oxygen concentration $[M/L^3]$, *Z* is river depth [L], and $k_{o_2}$ is *k* for oxygen at some temperature. When viewing equations 1 and 2 in aggregate, it is clear that *k* plays a critical role in two of the core, fundamental biogeochemical processes that occur in river networks. Both are reliant on accurate and high-fidelity assessments of *k*, as well as a rigorous understanding of the physical mechanisms driving *k*.
$$\mathbf{(2)} \frac{dDO}{dt}=\frac{M}{Z}+\frac{k_{o_2}}{Z}(DO_{air}-DO_{water})$$
Researchers use a range of methods to measure *k*, from floating domes to eddy covariance towers to measure at a specific point in the river. They also inject tracer gases to estimate reach-scale gas exchange [@hallGasExchangeStreams2020]. These approaches are ideal but are infeasible when working in ungauged rivers or at the river network scale across thousands of rivers. So, predictive models are used to estimate *k* from readily available river geomorphology [e.g. @oconnorMechanismReaerationNatural1958; @palumboAssessingPerformanceReaeration2014; @raymondScalingGasTransfer2012a]. These models often, though not always, predict $k_{600}$ or *k* that has been normalized to a Schmidt number of 600 ($CO_2$ at 20 degrees Celsius or $O_2$ at 17.5 degrees).

@@

**I'm struggling with this paragraph, and also don't know how to write it diplomatically**
We implement this model within BIKER (using 48 as a reasonable value for $\alpha$ given both forms of equation 3) for a few reasons. First, @wangPhysicallyBasedScaling2021 suggest that an $\alpha$ of approximately 48 is uniform across rivers and streams, regardless of the training data used. This is encouraging for use within BIKER as BIKER is specifically designed to be as river-agnostic as possible and so we sought as generalized an equation as possible. Second, it yields a simple linear relationship across all rivers, regardless of their size or steepeness. This is not true of other *k* models, for example those based on *eD* (section 1). The relationship between *eD* and *k* is not necessairly linear, with reasonable performance from both linear and power-law models and generally poorer predictive performance for small *k* values [@raymondScalingGasTransfer2012a]. Further, recent work has shown that *k* does not scale with *eD* via the same statistical parameters across all river geomorphologies [@ulsethDistinctAirWater2019a]. We further explored the relationship presented by [@ulsethDistinctAirWater2019a] in the SWOT context, finding that the $k \sim eD$ relationship fundamentally breaks down in rivers with low *eD*, which corresponds to nearly all rivers that SWOT will observe (Figure S2). Therefore, an *eD* based predictive model is less useful in SWOT rivers. Third, we sought to explicitly predict $k_{o_2}$, as its use extends beyond constraining gas fluxes from rivers and into parameterizing river metabolism models (equation 2). We stress that most other predictive models for *k* could be implemented within BIKER. However, our goal of using BIKER on any river that is observable by SWOT necessitates that we pick an equation that is applicable across as many rivers as possible. Future work should look at using a river-specific *k* equation to pseudo-calibrate BIKER to a specific river, as well as the influence of wind-induced gas exchange (see section 5.2).

@@

With the likelihood function (section 3.2.1), prior distributions (section 3.2.2), and the $\sigma_{k_{o_2}}^2$ parameter described (section 3.2.2), a joint posterior distribution conditional on the SWOT observations is specified. To approximate this distribution, we use a Markov Chain Monte Carlo (MCMC) algorithm implemented using the Stan probabilistic programming language. Specifically, Stan uses a Hamiltonian Monte Carlo sampler which reduces computation time relative to other sampling algorithms [@hagemannBAMBayesianAMHGManning2017a].

###3.3 Validation setup
We validate BIKER on the aforementioned 49 SWOT-simulated rivers (section 2) using daily observed $R_h$ and observed $S_h$. We also re-validate BIKER on the 17 rivers with the SWOT error model corrupting the SWOT-observables *W* and *H*. However, regardless of the validation data or SWOT error budget used, we do not have observed $k_{o_2}$ data for these rivers, and to our knowledge no field dataset of $k_{o_2}$ exists in the type of temporal and spatial frequency that SWOT (and therefore the BIKER algorithm) provides. Further, we are principally interested in BIKER's ability to reproduce $k_{o_2}$ from SWOT observations and are less concerned with the actual accuracy of the predictive model itself, which can be validated using existing datasets as has been previously done [e.g. @raymondScalingGasTransfer2012a; @ulsethDistinctAirWater2019a; @wangPhysicallyBasedScaling2021]. Therefore, we take the model outlined in section 3.1 and use that to calculate the observed $k_{o_2}$ that BIKER is validated against. This is done using equation 8, where $R_{h, obs} [L/T]$ is observed river channel area divided by observed channel width. This means that $k_{o_2, obs}$ relies on in situ river shear velocity, while BIKER is predicting shear velocity from *W* and $S_h$ alone. Therefore, we are directly exploring BIKER's ability to infer observed $U_\star$ and $k_{o_2}$ from *W* and *H* alone. It also means that, for a fair validation scheme, $\sigma_{k_{o_2}}^2$ must be set to reflect only error from our assumptions about $R_h$ and not the parameter uncertainity inherent in $\alpha$ (the 48 in equation 8). Thus, $\sigma_{k_{o_2}}^2$ is set to 0.20 for this validation as suggested in section 3.2. However once SWOT launches, it should reflect the total uncertainty calculated in section 3.2 from both sources.

@@

###3.4 Upscaling to CO2 evasion and bulk carbon efflux
It is one thing to accurately predict $k_{o_2}$, but researchers are often most interested in the actual gas fluxes from rivers and ultimately the carbon emitted from river to atmosphere. Therefore, we explore 1) the ability of BIKER to reproduce $CO_2$ evasion fluxes ($FCO_2$ $[M/L^2T]$) from these 49 rivers, and 2) the sensitivity of the carbon efflux $[M/T]$ due to the chosen method to approximate $R_h$ and ultimately $U_\star$.

First, we calculate $FCO_2$. To do this, we pair the biweekly 26 $CO_2$ samples (section 2) with every 11th SWOT observation by date, ignoring the timesteps beyond 26 (only ~15% of the SWOT observations were ignored here and we deemed this acceptable). We chose to sample every 11 days as this is the average sampling resolution for SWOT and the $CO_2$ data is approximately the same at a 14 day resolution. Not all of the SWOT rivers have observations for a full year, and when simulation dates were not available they were assumed to start on January 1st. We also pair the modeled $k_{CO_2}$ values (obtained from $k_{o_2}$ and equation 9) with these water-side $CO_2$ concentrations and water temperatures. In equation 9, 530 is the Schmidt number for oxygen at 20 degrees Celsius (which is the temperature at which BIKER predicts *k*) and *Sc* is the Schmidt number at one's desired temperature for some dissolved gas. This was calculated following @raymondScalingGasTransfer2012a and @wanninkhofRelationshipWindSpeed1992a. Atmospheric $CO_2$ was assumed 390 uatm. $FCO_2$ validation was performed using the same metrics as $k_{o_2}$ validation (Table 1).

@@

Next, we estimate bulk carbon efflux using five models for average channel depth (used to calculate $U_\star$):  BIKER's posterior means and four gauge-based HG models previously used for $CO_2$ upscaling [@horgbyUnexpectedLargeEvasion2019; @lauerwaldSpatialPatternsCO22015; @raymondGlobalCarbonDioxide2013a; @raymondScalingGasTransfer2012a]. See Table S1 for their definitions. It is worth stressing that the @lauerwaldSpatialPatternsCO22015 model is actually one of two components of the @raymondGlobalCarbonDioxide2013a model, and so was developed by the latter's authors. The names used here refer to the specific implementation used in each study. The @horgbyUnexpectedLargeEvasion2019 was developed explicitly for steep mountain streams, of which SWOT will not sample. We include it here to explore just how sensitive efflux estimates are to the HG model that is employed, noting that it will likely perform poorly. For this study, all HG methods use the in-situ discharge record while BIKER does not. This allows us to assess whether BIKER's estimates (wholly ungauged) are comparable to gauged methods (all four HG models). Finally, we express the bulk carbon efflux as the average mass flow rate of carbon (via $CO_2$ evasion) per year from the 49 rivers after accounting for total river surface area.
